---
title: 'Harmonizing the Domain 1 Marine Protected Area: new/updated krill and predators
  data to redefine boundaries and integrate seasonality in the MPA design'
author: "Lucas Krüger"
date: "2024-05-14"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This code uses the previous layers used as conservation objectives in the development of the design of the Domain 1 Marine Protected Area (D1MPA) and adds new data available from openly available peer-reviewed literature to redefine the boundaries of General Protection Zones and use seasonality of predator distribution to introduce temporal variability in the proposition of the MPA.

The code follows up from the previously standardized tracking data and standardized 2018 layers, files 01_Raw_Tracking_Data_Processment.R and 02_Krill_and_Predator_Layers_2018_rasterization.R


First, use the 'trip' package to regularize tracking data and calcualte cumulative area usage in 3-months periods (JFM, AMJ,JAS and OND)

```{r}
library(lubridate)
library(raster)
library(dplyr)
library(adehabitatHR)
library(adehabitatLT)
library(sp)
library(trip)
library(ggplot2)
library(patchwork)
library(terra)
library(reshape2)

mask <- raster("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/D1_Mask/D1_Mask.tif") # mask

plot(mask)

EPSG6932<-crs(mask) # extract the coordinate spatial system from the object

EPSG4326<-CRS("+proj=longlat +datum=WGS84 +no_defs")

D1<-shapefile("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Shapefiles/D1_6932.shp")


###----------Humpback Whale----------

hbw.ld2<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/Whale_Humpback.csv")

hbw.ld2$quarter<-quarter(hbw.ld2$date)

hbw.ld2$idy<-paste(hbw.ld2$id,year(hbw.ld2$date),month(hbw.ld2$date),week(hbw.ld2$date))

hbw.q1<-subset(hbw.ld2,quarter=="1")
hbw.q2<-subset(hbw.ld2,quarter=="2")


# JFM
d <- data.frame(x=hbw.q1$x,y=hbw.q1$y, tms=hbw.q1$date, id=hbw.q1$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/31.45

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Whale_Humpback_JFM.tif", overwrite=TRUE)


# AMJ 

d <- data.frame(x=hbw.q2$x,y=hbw.q2$y, tms=hbw.q2$date, id=hbw.q2$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/35.7

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Whale_Humpback_AMJ.tif", overwrite=TRUE)

###----------Minke WHale--------------
amw.ld2<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/Whale_AntarcticMinke.csv")

amw.q1<-subset(amw.ld2,quarter=="1")
amw.q2<-subset(amw.ld2,quarter=="2")


# JFM
d <- data.frame(x=amw.q1$x,y=amw.q1$y, tms=amw.q1$date, id=amw.q1$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/0.49

plot(ym)


writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Whale_AntMinke_JFM.tif", overwrite=TRUE)


# AMJ 

d <- data.frame(x=amw.q2$x,y=amw.q2$y, tms=amw.q2$date, id=amw.q2$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/0.20

plot(ym)


writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Whale_AntMinke_AMJ.tif", overwrite=TRUE)


#### -------------Crabeater Seal-----------------
crab.ld2<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/Seal_Crabeater.csv")

crab.q1<-subset(crab.ld2,quarter=="1")
crab.q2<-subset(crab.ld2,quarter=="2")
crab.q3<-subset(crab.ld2,quarter=="3")
crab.q4<-subset(crab.ld2,quarter=="4")


# amj

d <- data.frame(x=crab.q2$x,y=crab.q2$y, tms=crab.q2$date, id=crab.q2$id)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/71.221

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Crabeater_AMJ.tif", overwrite=TRUE)


# jas


d <- data.frame(x=crab.q3$x,y=crab.q3$y, tms=crab.q3$date, id=crab.q3$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/77.44

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Crabeater_JAS.tif", overwrite=TRUE)

#ond

d <- data.frame(x=crab.q4$x,y=crab.q4$y, tms=crab.q4$date, id=crab.q4$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/31.11

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Crabeater_OND.tif", overwrite=TRUE)


###------Antarctic Fur Seal----------------

afs.ld2<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/Seal_Fur.csv")

afs.ld2$quarter<-quarter(afs.ld2$date)
afs.q1<-subset(afs.ld2,quarter=="1")
afs.q2<-na.omit(subset(afs.ld2,quarter=="2"))
afs.q3<-subset(afs.ld2,quarter=="3")
afs.q4<-subset(afs.ld2,quarter=="4")

#jfm

d <- data.frame(x=afs.q1$x,y=afs.q1$y, tms=afs.q1$date, id=afs.q1$IDy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24

max(y)

ym<-y/17.7

plot(ym)


writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Fur_JFM.tif", overwrite=TRUE)

# amj

d <- data.frame(x=afs.q2$x,y=afs.q2$y, tms=afs.q2$date, id=afs.q2$IDy)
tr <- trip(d,correct_all = T)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24
y<-y+mask
max(y)

ym<-y/20.94

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Fur_AMJ.tif", overwrite=TRUE)


# jas
afs.q3$IDy<-paste(afs.q3$id,year(afs.q3$date),week(afs.q3$date))
d <- data.frame(x=afs.q3$x,y=afs.q3$y, tms=afs.q3$date, id=afs.q3$IDy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-cover(tg,mask)/(3600)/24
y<-y+mask
max(y)

ym<-y/7.24

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Fur_JAS.tif", overwrite=TRUE)

#ond

afs.q4$IDy<-paste(afs.q4$id,year(afs.q4$date),week(afs.q4$date))
d <- data.frame(x=afs.q4$x,y=afs.q4$y, tms=afs.q4$date, id=afs.q4$IDy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/4.73

plot(ym)

writeRaster(y,"C:/D1MPA trimestral/Trip/Seal_Fur_OND.tif", overwrite=TRUE)


###-----------Elephant Seal----------------

soes.ld2<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/Seal_Elephant.csv")

soes.ld2$quarter<-quarter(soes.ld2$date)
summary(as.factor(soes.ld2$quarter))

soes.q1<-subset(soes.ld2,quarter=="1")
soes.q2<-subset(soes.ld2,quarter=="2")
soes.q3<-subset(soes.ld2,quarter=="3")
soes.q4<-subset(soes.ld2,quarter=="4")


#jfm

d <- data.frame(x=soes.q1$x,y=soes.q1$y, tms=soes.q1$date, id=soes.q1$idy)
tr <- trip(d)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/156.1

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Elephant_JFM.tif", overwrite=TRUE)

# amj

dq2 <- data.frame(x=soes.q2$x,y=soes.q2$y, tms=soes.q2$date, id=soes.q2$idy)
tr2 <- trip(dq2,correct_all = T)
tg2 <- rasterize(tr2, mask, field = "tms") # in seconds

y2<-(cover(tg2,mask)/(3600)/24)+mask

max(y2)

ym2<-y2/108.01

plot(ym2)

writeRaster(ym2,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Elephant_AMJ.tif", overwrite=TRUE)

# jas

d3 <- data.frame(x=soes.q3$x,y=soes.q3$y, tms=soes.q3$date, id=soes.q3$idy)
tr3 <- trip(d3)
tg3 <- rasterize(tr3, mask, field = "tms") # in seconds

y3<-(cover(tg3,mask)/(3600)/24)+mask

max(y3)

ym3<-y3/77.001

plot(ym3)

writeRaster(ym3,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Elephant_JAS.tif", overwrite=TRUE)

#ond
d4 <- data.frame(x=soes.q4$x,y=soes.q4$y, tms=soes.q4$date, id=soes.q4$idy)
tr4 <- trip(d4)

tg4 <- rasterize(tr4, mask, field = "tms") # in seconds

y4<-(cover(tg4,mask)/(3600)/24)+mask

max(y4)

ym4<-y4/367.01

plot(ym4)

writeRaster(ym4,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Seal_Elephant_OND.tif", overwrite=TRUE)


###-------Penguins------------

pengm<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Processed_Tracking_Data/penguins.csv")

pengm$timestamp<-as.POSIXct(strptime(pengm$timestamp,
                                     format="%Y-%m-%d %H:%M:%S", tz="GMT"))
pengm$month<-month(pengm$timestamp)
pengm$id<-paste(pengm$Spp,pengm$id,pengm$month)

#pengm<-subset(pengm,month<9)

pengm2<-plyr::ddply(pengm, c("id"), summarise,
                    N=length(lon)) 

pengms<-merge(pengm,pengm2,by="id")

pengms<-subset(pengms,N.y>5)

adp<-subset(pengms,Spp=="ADPE")
chp<-subset(pengms,Spp=="CHPE")
gep<-subset(pengms,Spp=="GEPE")

adp<-subset(adp,lon>(-75))

gep<-subset(gep,lon<(-50) & lon>(-67))
gep<-subset(gep,lat<(-61.5))

adsp<-SpatialPointsDataFrame(adp[6:7],adp,proj4string = EPSG4326)
chsp<-SpatialPointsDataFrame(chp[6:7],chp,proj4string = EPSG4326)
gesp<-SpatialPointsDataFrame(gep[6:7],gep,proj4string = EPSG4326)

adspW<-sp::spTransform(adsp,CRS("+proj=laea +lat_0=-90 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"))
chspW<-sp::spTransform(chsp,CRS("+proj=laea +lat_0=-90 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"))
gespW<-sp::spTransform(gesp,CRS("+proj=laea +lat_0=-90 +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"))

plot(adspW)
plot(chspW)
plot(gespW)

### --------gentoo penguin----------------

gepW<-raster::intersect(gespW,D1)

geps<-data.frame(gepW)

head(geps)
geps<- geps[!duplicated(geps[, c("timestamp","id")]), ]

gep.lt<-as.ltraj(xy=geps[18:19],date=geps$timestamp,id=geps$id,proj4string = EPSG6932)

gep.ld<-ld(gep.lt)

ggplot(gep.ld,aes(dist))+geom_histogram()

gep.ld2<-subset(gep.ld,dist<50000) #eliminate positions based on consecutive points distance 

gep.ld2$year<-year(gep.ld2$date)

gep.ld2$quarter<-quarter(gep.ld2$date)
gep.q1<-subset(gep.ld2,quarter=="1")
gep.q2<-subset(gep.ld2,quarter=="2")
gep.q3<-subset(gep.ld2,quarter=="3")
gep.q4<-subset(gep.ld2,quarter=="4")

plot(gep.q2$x,gep.q2$y)

# amj

d <- data.frame(x=gep.q2$x,y=gep.q2$y, tms=gep.q2$date, id=gep.q2$id)
tr <- trip(d,correct_all = T)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/8.71

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Penguin_Gentoo_AMJ.tif", overwrite=TRUE)

# jas

d3 <- data.frame(x=gep.q3$x,y=gep.q3$y, tms=gep.q3$date, id=gep.q3$id)
tr3 <- trip(d3,correct_all = T)

tg3 <- rasterize(tr3, mask, field = "tms") # in seconds

y3<-(cover(tg3,mask)/(3600)/24)+mask

max(y3)

ym3<-y3/1.14

plot(ym3)

writeRaster(ym3,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Penguin_Gentoo_JAS.tif", overwrite=TRUE)

### -----chinstrap-------

summary(as.factor(chp$Col))

chspW<-raster::intersect(chspW,D1)

chps<-data.frame(chspW)

chps<- chps[!duplicated(chps[, c("timestamp","id")]), ]

head(chps)

chp.lt<-as.ltraj(xy=chps[18:19],date=chps$timestamp,id=chps$id,proj4string = EPSG6932)

chp.ld<-ld(chp.lt)

ggplot(chp.ld,aes(dist))+geom_histogram()


chp.ld2<-subset(chp.ld,dist<50000) #eliminate positions based on consecutive points distance 

chp.ld2$quarter<-quarter(chp.ld2$date)
chp.q1<-subset(chp.ld2,quarter=="1")
chp.q2<-subset(chp.ld2,quarter=="2")
chp.q3<-subset(chp.ld2,quarter=="3")
chp.q4<-subset(chp.ld2,quarter=="4")

# amj

d <- data.frame(x=chp.q2$x,y=chp.q2$y, tms=chp.q2$date, id=chp.q2$id)

tr <- trip(d,correct_all = T)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/1.153

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Penguin_Chinstrap_AMJ.tif", overwrite=TRUE)

# jas

d <- data.frame(x=chp.q3$x,y=chp.q3$y, tms=chp.q3$date, id=chp.q3$id)
tr <- trip(d,correct_all = T)

tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/0.15

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Penguin_Chinstrap_JAS.tif", overwrite=TRUE)

##----------adelie penguin--------------

summary(as.factor(adp$Col))

adspW<-raster::intersect(adspW,D1)

adps<-data.frame(adspW)


adps<- adps[!duplicated(adps[, c("timestamp","id")]), ]

head(adps)

adp.lt<-as.ltraj(xy=adps[18:19],date=adps$timestamp,id=adps$id,proj4string = EPSG6932)

adp.ld<-ld(adp.lt)

ggplot(adp.ld,aes(x,y))+geom_point()

ggplot(adp.ld,aes(dist))+geom_histogram()

adp.ld2<-subset(adp.ld,dist<50000) #eliminate positions based on consecutive points distance 

adp.ld2$month<-month(adp.ld2$date)

adspM<-subset(adp.ld2,month=="2"|month=="3"|month=="4"|month=="1")


d <- data.frame(x=adspM$x,y=adspM$y, tms=adspM$date, id=adspM$id)
tr <- trip(d,correct_all = T)
tg <- rasterize(tr, mask, field = "tms") # in seconds

y<-(cover(tg,mask)/(3600)/24)+mask

max(y)

ym<-y/10.1

plot(ym)

writeRaster(ym,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/Penguin_Adelie_Mig.tif", overwrite=TRUE)


```

## 

Now join all the predators layers to generate a grid file that represents the number of layers, which is one of the criteria used to redefine GPZ.After generating those files, vertices for the new boundaries were drawn in ArcMap, and a .csv file was generated.

```{r}

###------------- rescale data to binary ----------------

# load files

# Set the directory where TIF files are located
tif_directory <- "C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Predators_rescaled_area_usage/"

# List all TIF files in the directory
tif_files <- list.files(tif_directory, pattern = ".tif$", full.names = T)

# Load all TIF files into a raster stack
raster_stack <- terra::rast(tif_files)

# Define the number of intervals
num_intervals <- 2


for (i in 1:nlyr(raster_stack)) {
  # Extract the ith layer
  raster_layer <- raster_stack[[i]]
  
  # Calculate the 0.05 quantile
  quantile_005 <- quantile(values(raster_layer), probs = 0.95,na.rm=T)
  
  # Reclassify the raster layer
  reclassified_layer <- terra::ifel(raster_layer > quantile_005, 1, 0)
  
  # Replace the original layer with the reclassified layer in the raster stack
  raster_stack[[i]] <- reclassified_layer
}


# Plot the reclassified raster stack
plot(raster_stack)

# load additional binary files 

# Set the directory where TIF files are located
tif_directory2 <- "C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Non-tracking_layers/"

# List all TIF files in the directory
tif_files2 <- list.files(tif_directory2, pattern = ".tif$", full.names = T)


# Load all TIF files into a raster stack
raster_stack2 <- terra::rast(tif_files2)
plot(raster_stack2)


#stack both stacks together
rasters<-c(raster_stack,raster_stack2)

rsum<-sum(rasters) # sum all stacks

plot(rsum)

max(rsum)

writeRaster(rsum,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/UpDated_Layers_Sum.tif",overwrite=T)


### -------2018 layers------------

# load additional binary files 

# Set the directory where TIF files are located
tif_directory3 <- "C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Layers2018/geotiff"

# List all TIF files in the directory
tif_files3 <- list.files(tif_directory3, pattern = ".tif$", full.names = T)

tif_files3

# Load all TIF files into a raster stack

raster_stack3 <- terra::rast(tif_files3)

plot(raster_stack3)

layers_2018<-sum(raster_stack3)

writeRaster(layers_2018,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Layers2018_Sum.tif",overwrite=T)

#stack all stacks together
rasters<-c(raster_stack,raster_stack2,raster_stack3)

rsum2<-sum(rasters) # sum all stacks

plot(rsum2)

max(rsum2)


writeRaster(rsum2,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/AllLayers_Sum.tif",overwrite=T)


```

Next, with the vertices files, generate a polygon

```{r}

library(raster)
library(terra)
library(sf)

Peninsula<-readRDS("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/VerticesFiles/Peninsula.Rds") # land shapefile downloaded from the Antarctica Digital data base ('Gerrish, L., Ireland, L., Fretwell, P., & Cooper, P. (2023). High resolution vector polygons of the Antarctic coastline (7.7) [Data set]. UK Polar Data Centre, Natural Environment Research Council, UK Research & Innovation. https://doi.org/10.5285/0be5339c-9d35-44c9-a10f-da4b5356840b') and subsetted to match Domain 1 limits

crs<-crs(Peninsula)  # coordinate system, South Polar Lambert Azimuthal Equal Area, ESRI 102020

D1MPA<-read.csv("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/VerticesFiles/D1MPA_Vertices_2024.csv") # D1MPA vertices

wap = st_as_sf(Peninsula)


xys = st_as_sf(D1MPA, coords=c("Lon_ESRI_102020","Lat_ESRI_102020"),crs=crs) # convert to a simple feature object

polys = xys %>% 
  dplyr::group_by(Name) %>% 
  dplyr::summarise(geometry = st_combine(geometry)) %>%
  st_cast("POLYGON")

plot(polys)


### now eliminate the land from the MPA

diffPoly <- st_difference(polys, st_union(wap))  



diffPoly$Zone[diffPoly$Name=="EI" | diffPoly$Name=="SSIW" |
                diffPoly$Name=="SWAP" |  diffPoly$Name=="NWAP"|
                diffPoly$Name=="SOI" ]<-"GPZ"

diffPoly$Zone[is.na(diffPoly$Zone)]<-"SPZ"

plot(diffPoly)

# save as shapefile to use on other GIS
st_write(diffPoly, dsn = "C:/D1MPA_2024_Harmony/D1MPA_Harmonized/GPZ-SPZ",
         layer="D1MPA_Vertices_2024", driver="ESRI Shapefile",
         append=FALSE)

#or save as Rds

saveRDS(diffPoly,"C:/D1MPA_2024_Harmony/D1MPA_Harmonized/GPZ-SPZ/VerticesD1MPA_Vertices_2023.Rds")



```

And finally, calculate layers coverage (based on the overlap with fishable areas, see main document) to assign seasonal protection zones to those layers whose GPZ coverage was below the 50%.

```{r}


###--- coverage --------

### load subareas and strata files

strata<-shapefile("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Shapefiles/Subareas_Strata.shp")
plot(strata)

summary(as.factor(strata$SubArea))

gpz<-shapefile("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/GPZ-SPZ/D1MPA_Vertices_2024.shp")
head(gpz)
summary(as.factor(gpz$Name))
summary(as.factor(gpz$Zone))
plot(gpz)

# preferred fishable area

fa<-shapefile("C:/D1MPA_2024_Harmony/D1MPA_Harmonized/Shapefiles/Fishable_Area.shp")
plot(fa)

### spatial points

spt<-shapefile("C:/D1MPA trimestral/D1MPA Fishable Area/Shapefiles/Grid_Points_D1.shp")

head(spt)

stfa<- (over(x = spt, y = fa))

summary(as.factor(stfa$Fishable))

stfa<-SpatialPointsDataFrame(coordinates(spt),stfa[4],proj4string = crs(spt))

#-----strata coverage -----------

st.points<- (over(x = stfa, y = strata))

summary(as.factor(st.points$SubArea))

stratas<-data.frame(st.points[4],stfa[1], coordinates(spt))

stgrid<-SpatialPointsDataFrame(coordinates(spt),stratas)

#plot(stgrid)

stvect<-vect(stgrid)

head(stvect)

summary(as.factor(stvect$SubArea))

ext.st<-terra::extract(rasters, stvect)

head(ext.st)

st.df <- data.frame(FA=stvect$Fishable,SA=stvect$SubArea,ext.st[2:54])

head(st.df)

summary(as.factor(st.df$SA))
summary(as.factor(st.df$FA))


###--------- gpz covergae-----

gpz.points<- (over(x = stfa, y = gpz))

gpzs<-data.frame(gpz.points)
head(gpzs)
summary(as.factor(gpzs$Name))

summary(as.factor(gpzs$Zone))

sfdf<-data.frame(st.df[1:2],gpzs)

head(sfdf)

gpzgrid<-SpatialPointsDataFrame(coordinates(stfa),sfdf)

gpzvect<-vect(gpzgrid)

head(gpzvect)

ext.gpz<-terra::extract(rasters, gpzvect)

gpz.df <- data.frame(sfdf[1:4],ext.gpz[2:54])

ggplot(gpz.df,aes(SA,fill=Zone))+geom_bar(stat = "count")

#gpz.df$SA[is.na(gpz.df$SA)]<-"Outside"

summary(as.factor(gpz.df$FA))
summary(as.factor(gpz.df$Name))


###-----coverage calcualtion ---------

total<- data.frame(st.df[3:55]) %>%
  #group_by(FA) %>%
  summarise_all(sum, na.rm = TRUE)

totalF<- data.frame(st.df[1],st.df[3:55]) %>%
  group_by(FA) %>%
  summarise_all(sum, na.rm = TRUE)

totalF<-subset(totalF,FA=="Yes")


SAO <- st.df[2:55] %>%
  group_by(SA) %>%
  summarise_all(sum, na.rm = TRUE)

SAF <- st.df %>%
  group_by(FA,SA) %>%
  summarise_all(sum, na.rm = TRUE)

SAF<-subset(SAF,FA=="Yes")
SAO
SAF


head(gpz.df)


gpz_only<-subset(gpz.df,Zone=="GPZ")

summary(as.factor(gpz_only$SA))

GPZ <- data.frame(gpz_only[2],gpz_only[5:57]) %>%
  group_by(SA) %>%
  summarise_all(sum, na.rm = TRUE)

GPF <- data.frame(gpz_only[1:2],gpz_only[5:57]) %>%
  group_by(FA,SA) %>%
  summarise_all(sum, na.rm = TRUE)

GPF<-subset(GPF,FA=="Yes")

totm<-melt(total)
tofm<-melt(totalF)

saom<-melt(SAO,id.vars=c("SA"))
safm<-melt(SAF,id.vars=c("FA","SA"))

gpzm<-na.omit(melt(GPZ,id.vars=c("SA")))
gpfm<-na.omit(melt(GPF,id.vars=c("FA","SA")))

names(totm)[names(totm) == "value"] <- "range"
names(saom)[names(saom) == "value"] <- "SA.cover"
names(gpzm)[names(gpzm) == "value"] <- "GPZ.cover"

names(tofm)[names(tofm) == "value"] <- "rangeF"
names(safm)[names(safm) == "value"] <- "SA.coverF"
names(gpfm)[names(gpfm) == "value"] <- "GPZ.coverF"

head(gpfm)

head(safm)

head(tofm)

gpz.st<-merge(gpzm,saom,all=T)

alldf<-merge(gpz.st,totm,all=T)
head(alldf)

alldff1<-merge(alldf,tofm)
alldff2<-merge(alldff1,safm)
alldf<-merge(alldff2,gpfm)

head(alldf)

names(alldf)[names(alldf) == "variable"] <- "layers"

alldf$sa.fa.prop<-alldf$SA.coverF/alldf$SA.cover

alldf$sa.fa.prop[is.na(alldf$sa.fa.prop)]<-0


### if a layer has more than 75% of its occurrence within fishable area, it is a priority


alldf$priority<-ifelse(alldf$sa.fa.prop>0.75,"Priority","Non-priority")

alldf$GPZ.prop<-alldf$GPZ.cover/alldf$SA.cover

alldf$GPZ.prop[is.na(alldf$GPZ.prop)]<-0



head(alldf)

# specify layers. If it starts with K they are krill for the updated layers
# if O, they are layers from 2018, except krill that is "o"
# if other, they are updated predator layers

alldf$type<-substring(alldf$layers,first=1,last=1)
summary(as.factor(alldf$layers))
summary(as.factor(alldf$type))

alldf$types[alldf$type=="K"]<-"Krill updated"
alldf$types[alldf$type=="O"]<-"layers 2018"
alldf$types[is.na(alldf$types)]<-"predators updated"

alldf$types[alldf$layers=="obj_5b_Esuperba"]<-"Krill updated"

summary(as.factor(alldf$types))

layers_2018<-subset(alldf,types=="layers 2018")
predators<-subset(alldf,types=="predators updated")

krill<-subset(alldf,types=="Krill updated" )


gpzS<-subset(predators,layers!="Penguin_Gentoo_AMJ" & layers!="Penguin_Gentoo_JAS" &
               layers!="Penguin_Chinstrap_AMJ" & layers!="Penguin_Chinstrap_JAS"&
               layers!="Seal_Crabeater_AMJ"&layers!="Seal_Crabeater_JAS"&
               layers!="Seal_Elephant_AMJ"&layers!="Seal_Elephant_JAS"&
               layers!="Seal_Fur_AMJ"&layers!="Seal_Fur_JAS"&
               layers!="Whale_Humpback_AMJ"&layers!="Whale_AntMinke_AMJ"&
               layers!="Penguin_Emperor_Cols" & layers!="obj_5b_Esuperba" & 
               layers!="AFS_BS_SOI" )
#winter layers


gpzW<-subset(predators,layers=="Penguin_Gentoo_AMJ" | layers=="Penguin_Gentoo_JAS" |
               layers=="Penguin_Chinstrap_AMJ" | layers=="Penguin_Chinstrap_JAS"|
               layers=="Seal_Crabeater_AMJ"|layers=="Seal_Crabeater_JAS"|
               layers=="Seal_Elephant_AMJ"|layers=="Seal_Elephant_JAS"|
               layers=="Seal_Fur_AMJ"|layers=="Seal_Fur_JAS"|
               layers=="Whale_Humpback_AMJ"|layers=="Whale_AntMinke_AMJ"|
               layers=="Penguin_Emperor_Cols")


# 2018 layers 

(ggplot(layers_2018,aes(layers,GPZ.prop,colour=priority))+
    geom_segment( aes(x=layers, xend=layers, y=0.5, yend=GPZ.prop),
                  linetype="dotted") +
    geom_point(size=2)+
    geom_hline(yintercept=c(0.5))+
    facet_wrap(SA~.)+coord_flip()+
    scale_colour_manual(values=c("blue","red"))+
    ggtitle(label="")+theme_bw()+
    xlab("layers")+ylab("Proportion of GPZ coverage")+
    ggtitle(label="Layers 2018"))

### layers 2024 

(ggplot((krill),aes(layers,GPZ.prop,colour=priority))+
    geom_segment( aes(x=layers, xend=layers, y=0.2, yend=GPZ.prop),
                  linetype="dotted") +
    geom_point(size=2)+
    geom_hline(yintercept=c(0.2))+
    facet_wrap(SA~.)+coord_flip()+
    scale_colour_manual(values=c("blue","red"))+
    ggtitle(label="")+theme_bw()+
    xlab("Summer layers")+ylab("Proportion of GPZ coverage")+
    ggtitle(label="a. Layers 2024 Krill"))/
  
  # summer layers
  
  (ggplot((gpzS),aes(layers,GPZ.prop,colour=priority))+
     geom_segment( aes(x=layers, xend=layers, y=0.5, yend=GPZ.prop),
                   linetype="dotted") +
     geom_point(size=2)+
     geom_hline(yintercept=c(0.5))+
     facet_wrap(SA~.)+coord_flip()+
     scale_colour_manual(values=c("blue","red"))+
     ggtitle(label="")+theme_bw()+
     xlab("Summer layers")+ylab("Proportion of GPZ coverage")+
     ggtitle(label="b. Layers 2024 October to March"))/
  
  # a part of fish habitat is inside SOISS MPA. That might increase coverage
  
  
  
  #Winter layers
  
  (ggplot((gpzW),aes(layers,GPZ.prop,colour=priority))+
     geom_segment( aes(x=layers, xend=layers, y=0.5, yend=GPZ.prop),
                   linetype="dotted") +
     geom_point(size=2)+
     geom_hline(yintercept=c(0.5))+
     facet_wrap(SA~.)+coord_flip()+
     scale_colour_manual(values=c("blue","red"))+
     ggtitle(label="")+theme_bw()+
     xlab("Winter layers")+ylab("Proportion of GPZ coverage")+
     ggtitle(label="Layers 2024 April to September"))




###---------- now, do the same with the SPZ considered----------


###-----coverage calcualtion ---------

head(gpz.df)

gpz_spz<-subset(gpz.df,Zone=="GPZ"|Zone=="SPZ")

summary(as.factor(gpz_spz$SA))

SPZ <- data.frame(gpz_spz[2],gpz_spz[5:57]) %>%
  group_by(SA) %>%
  summarise_all(sum, na.rm = TRUE)

SPF <- data.frame(gpz_spz[1:2],gpz_spz[5:57]) %>%
  group_by(FA,SA) %>%
  summarise_all(sum, na.rm = TRUE)

SPF<-subset(SPF,FA=="Yes")

spzm<-na.omit(melt(SPZ,id.vars=c("SA")))
spfm<-na.omit(melt(SPF,id.vars=c("FA","SA")))
summary(as.factor(spzm$SA))
summary(as.factor(spfm$SA))

names(spzm)[names(spzm) == "value"] <- "GPZ.cover"

names(spfm)[names(spfm) == "value"] <- "GPZ.coverF"

head(spfm)

spz.st<-merge(spzm,saom,all=T)

all<-merge(spz.st,totm,all=T)
head(all)

allf1<-merge(all,tofm)
allf2<-merge(allf1,safm)
allf<-merge(allf2,spfm)

head(allf)

names(allf)[names(allf) == "variable"] <- "layers"

allf$sa.fa.prop<-allf$SA.coverF/allf$SA.cover

allf$sa.fa.prop[is.na(allf$sa.fa.prop)]<-0


### if a layer has more than 75% of its occurrence within fishable area, it is a priority


allf$priority<-ifelse(allf$sa.fa.prop>0.75,"Priority","Non-priority")

allf$GPZ.prop<-allf$GPZ.cover/allf$SA.cover

allf$GPZ.prop[is.na(allf$GPZ.prop)]<-0

head(allf)


# specify layers. If it starts with X they are krill for the updated layers
# if O, they are layers from 2018
# if other, they are updated predator layers

allf$type<-substring(allf$layers,first=1,last=1)

summary(as.factor(allf$type))

allf$types[allf$type=="X"]<-"Krill updated"
allf$types[allf$type=="O"]<-"layers 2018"
allf$types[is.na(allf$types)]<-"predators updated"

allf$types[allf$layers=="obj_5b_Esuperba"]<-"Krill updated"

prior<-subset(allf,priority=="Priority")
head(prior)

krillSPZ<-subset(prior,types=="Krill updated" )
predatorsS<-subset(prior,types=="predators updated")

gpzS1<-subset(predatorsS,layers!="Penguin_Gentoo_AMJ" & layers!="Penguin_Gentoo_JAS" &
                 layers!="Penguin_Chinstrap_AMJ" & layers!="Penguin_Chinstrap_JAS"&
                 layers!="Seal_Crabeater_AMJ"&layers!="Seal_Crabeater_JAS"&
                 layers!="Seal_Elephant_AMJ"&layers!="Seal_Elephant_JAS"&
                 layers!="Seal_Fur_AMJ"&layers!="Seal_Fur_JAS"&
                 layers!="Whale_Humpback_AMJ"&layers!="Whale_AntMinke_AMJ"&
                 layers!="Penguin_Emperor_Cols" & layers!="obj_5b_Esuperba" & 
                 layers!="AFS_BS_SOI" & layers!="Seal_Crabeater_OND" )
head(gpzS)

gpzS0<-subset(gpzS,priority=="Priority")

gpzS0$unit<-c("GPZ")
gpzS1$unit<-c("GPZ + SPZ")

head(gpzS0)
head(gpzS1)

gpzspz2<-rbind(gpzS0,gpzS1)


### layers 2024 GPZ and GPZ+SPZ
summary(as.factor(gpzspz2$layers))

gpzspz2$layers<-factor(gpzspz2$layers,levels=c("bathy_150_500","bathy_0_150",
                                               "Whale_Humpback_JFM",
                                               "Whale_AntMinke_JFM","Seal_Elephant_JFM",
                                               "IBA_CHP", "IBA_GEP","IBA_ADP"))

(ggplot(na.omit(subset(gpzspz2,SA=="SA 48.1")),aes(layers,GPZ.prop,colour=unit))+
    geom_segment( aes(x=layers, xend=layers, y=0.5, yend=GPZ.prop),
                  linetype="dotted") +
    geom_point(size=2)+
    geom_hline(yintercept=c(0.5),linetype="dashed")+
    facet_wrap(SA~.)+coord_flip()+
    scale_colour_manual(values=c("blue","red"))+
    ggtitle(label="")+theme_bw()+
    xlab("Summer layers")+ylab("Proportion")+
    ggtitle(label=" layers GPZ + SPZ coverage"))


```
